*Prochlorococcus marinus* responses to light and oxygen

### 2024-05-24 Code/Manuscript_O2.Rmd
Added temperate niche table

Fig \@ref(fig:DNARepairFig) shows genes encoding DNA repair for *P.
marinus* strains. As expected, *P. marinus* MED4 possesses the largest,
most complete suite of genes encoding DNA repair enzymes, followed by
*P. marinus* MIT9313. Conversely, *P. marinus* SS120 demonstrates the
smallest genomic capacity for DNA repair. *Prochlorococcus marinus* MED4
and NATL2A were the only strains to possess a gene encoding
deoxyribodipyrimidine photolyase (Fig \@ref(fig:DNARepairFig) and
\@ref(fig:ProchlorococcusLightEnzymes)), which, in the presence of blue
light, is responsible for repairing DNA damaged by UV light
[@sancarStructureFunctionDNA2003]. *Prochlorococcus marinus* MED4 was
also the only strain to possess a gene encoding DNA repair ligase, which uses
ATP as a cofactor for DNA repair. The absence of genes encoding
deoxyribodipyrimidine photolyase and DNA repair ligase (ATP) in *P. marinus*
MIT9313 and *P. marinus* SS120 may, in part, explain why these two strains cannot tolerate growth under full [O~2~] and high light, found at the ocean
surface. Furthermore, the protective effect of lower [O~2~], allowing
these strains to grow at higher light, may relate in part to suppression
of DNA damage when generation of Reactive Oxygen Species is suppressed
at lower [O~2~]. NATL2A, a clade LLI, has been found near the ocean
surface during deep ocean mixing
[@malmstromTemporalDynamicsProchlorococcus2010]. Malmstrom *et al*.
[@malmstromTemporalDynamicsProchlorococcus2010] attributes NATL2A
tolerance to short exposures of high light to the presence of the genes
encoding photolyase, a gene found in HL clades. The presence of
deoxyribodipyrimidine photolyase and absence of DNA repair ligase (ATP)
supports why NATL2A tolerates limited exposure to high light and why
NATL2A is unable to fully repair damaged DNA. *Prochlorococcus* are
highly susceptible to hydrogen peroxide (H~2~O~2~) toxicity as they lack
genes which scavenge H~2~O~2~ molecules
[@morrisFacilitationRobustGrowth2008]. The small cell size of
*Prochlorococcus* allow the reactive oxygen species (ROS), H~2~O~2~, to
cross the cell membrane [@omarDiffusionalInteractionsMarine2022];
however, accumulation of extracellular H~2~O~2~ remains toxic to
*Prochlorococcus* [@morrisFacilitationRobustGrowth2008;
@morrisDependenceCyanobacteriumProchlorococcus2011].

@@

>>>>

Fig \@ref(fig:DNARepairFig) shows genes encoding DNA repair for *P.marinus* strains. As expected, *P. marinus* MED4 possesses the largest,
most complete suite of genes encoding DNA repair enzymes, followed by
*P. marinus* MIT9313. Conversely, *P. marinus* SS120 demonstrates the
smallest genomic capacity for DNA repair. *Prochlorococcus marinus* MED4
and NATL2A were the only strains to possess a gene encoding
deoxyribodipyrimidine photolyase (Fig \@ref(fig:DNARepairFig) and
\@ref(fig:ProchlorococcusLightEnzymes)), which, in the presence of blue
light, is responsible for repairing DNA damaged by UV light
[@sancarStructureFunctionDNA2003]. *Prochlorococcus marinus* MED4 was
also the only strain to possess a gene encoding DNA repair ligase, which uses
ATP as a cofactor for DNA repair. The absence of genes encoding
deoxyribodipyrimidine photolyase and DNA repair ligase (ATP) in *P. marinus*
MIT9313 and *P. marinus* SS120 may, in part, explain why these two strains cannot tolerate growth under full [O~2~] and high light, found at the ocean
surface. Furthermore, the protective effect of lower [O~2~], allowing
these strains to grow at higher light, may relate in part to suppression
of DNA damage when generation of Reactive Oxygen Species is suppressed
at lower [O~2~]. NATL2A, a clade LLI, has been found near the ocean
surface during deep ocean mixing
[@malmstromTemporalDynamicsProchlorococcus2010]. Malmstrom *et al*.
[@malmstromTemporalDynamicsProchlorococcus2010] attributes NATL2A
tolerance to short exposures of high light to the presence of the genes
encoding photolyase, a gene found in HL clades. The presence of
deoxyribodipyrimidine photolyase and absence of DNA repair ligase (ATP)
supports why NATL2A tolerates limited exposure to high light and why
NATL2A is unable to fully repair damaged DNA. *Prochlorococcus* are
highly susceptible to hydrogen peroxide (H~2~O~2~) toxicity as they lack
genes which scavenge H~2~O~2~ molecules
[@morrisFacilitationRobustGrowth2008]. The small cell size of
*Prochlorococcus* allow the reactive oxygen species (ROS), H~2~O~2~, to
cross the cell membrane [@omarDiffusionalInteractionsMarine2022];
however, accumulation of extracellular H~2~O~2~ remains toxic to
*Prochlorococcus* [@morrisFacilitationRobustGrowth2008;
@morrisDependenceCyanobacteriumProchlorococcus2011].

@@

```{r read temperate niche sheet, include=FALSE}

ProchloroNiches <- read_csv(file.path(DataIn, "Prochlorococcus_temperate_niches.csv"))

ProchloroNiches <- ProchloroNiches %>% 
  mutate(MED4 = case_when(MED4 == "Potential" ~ "+",
                          MED4 == "Excluded" ~ "-"),
         SS120 = case_when(SS120 == "Potential" ~ "+",
                          SS120 == "Excluded" ~ "-"),
         MIT9313 = case_when(MIT9313 == "Potential" ~ "+",
                          MIT9313 == "Excluded" ~ "-"))

```{r NichesTable,  fig.height = 8, fig.width = 6, echo = FALSE}

knitr::kable(ProchloroNiches,
             caption = "Potential temperate niches for *Prochlorococcus marinus* ecotypes defined by season, depth, and oxygen; Full denotes 250 µM of dissolved oxygen and OMZ denotes 25 µM of dissolved oxygen. Plus symbols (+) indicate a potential niche, minus symbols (-) not a potential niche.",
        align = "lllccc" )
```
